### 2023-11-14 manuscript.Rmd
Import the manuscript
This is a mostly clean conversion including all formulas and figures.
The WWW references are not fully satisfactory yet, but that could be
amended later. Currently, bookdown features (cross-references) are used.

---
title: Modeling HEPES buffer with a charge-balance approach
author: Troels Ring, Søren Risom Kristensen, Ivan Krylov, John C Nash.
bibliography: manuscript.bib
---

```{r, echo=FALSE}
# NOTE: bookdown features (e.g. cross-references) used below
# This requires the use of a bookdown output format, e.g.
# bookdown::pdf_document2 or bookdown::html_document2
```

Abstract
========

Introduction
============

It has been shown [@Ring2022] that $\mathrm{pH}$ in arbitrary fluids can be
understood based on the three fundamental principles of
electroneutrality, conservation of mass, and rules of dissociation as
devised from physical chemistry. It has also been shown [@Ring2022] that the
charge-balance approach allows definition of the $\mathrm{pH}$-dependent
buffer-capacity, and provides a definition of acid content as titratable
acidity, based on previous work by Kildeberg [@Kildeberg1983]. Recently
it has been reiterated [@Ring2023-charge] that it is absolutely necessary for this
modeling approach to take into account the strong ions, which are those
assumed to be completely dissociated at any $\mathrm{pH}$. In contrast to the
physico-chemical model of Stewart [@Stewart1978], the charge-balance
model (CBM, henceforth) has no separation between independent and
dependent variables. Under the CBM all that can be known about the
system is gathered in the expression of charge in an explicitly stated
chemical system. Consequently, if, as an example, $\mathrm{pH}$ is known and charge
is a known deterministic function of the chemical specification, inverse
inferences on unknown variables can, at least sometimes, be obtained by
root finding, since we know that the charge must be zero. In simple
systems, such as a single monovalent weak acid and strong ions, a single
unknown variable is explicitly estimable [@Ring2022]. The question
focused here is if CBM is rich enough to allow inverse inference in more
complex systems such as with multiprotic buffers. This question arose
under experiments with HEPES buffered sulfate containing fluids
[@Ring2023-acid] in which HEPES was modeled as a monovalent buffer with
pKa 7.5 [@Ring2010] since it was found that estimated $\mathrm{pH}$ was much too
low when observed $\mathrm{pH}$ was less than about 6. HEPES is a zwitterion with
formula 2-[4-(2-Hydroxyethyl)piperazin-1-yl]ethane-1-sulfonic acid and
structure as shown in \@ref(fig:hepes).

```{r hepes, fig.cap = '2-[4-(2-Hydroxyethyl)piperazin-1-yl]ethane-1-sulfonic acid', echo = FALSE}
knitr::include_graphics('HEPES.svg', auto_pdf = TRUE)
```

Wikipedia informs [@WP:HEPES] that HEPES has two pK values, one at 3 and
one at 7.5, but the reference given [@Johnson2014] in Wikipedia states
that "3" comes from the sulfonate group and 7.5 from one of the
piperazine nitrogens.  At $20^\circ\mathrm{C}$ Good et al [@Good1966]
has pKa 3 and 7.55 without detailing which groups are at work, while
another study [@AlGhobashy2014] mentions that 3 and 7.5 both belongs to
piperazine amines and informs that the sulfonate group is dissociated
over "almost the whole $\mathrm{pH}$ range".  Furthermore, is has been stated that
the nitrogen at the $\mathrm{OH}$ end of HEPES is protonated in methanol
and the nitrogen in the sulfonate end is protonated in water
[@Sledz2010].

Hence it was not clear how best to model the $\mathrm{pH}$-dependent
charge on HEPES to estimate $\mathrm{pH}$ based on known chemistry.
Furthermore, it was desired to make clear if the $\mathrm{p}K$ values
for HEPES could be elucidated based on measured $\mathrm{pH}$. Herein,
these questions will be approached by first developing a model for the
mean charge on HEPES as a function of $\mathrm{pH}$ and applying this to
proposed values of set of underlying pKs using data obtained previously
[@Ring2023-acid]. Next, it will be examined if data generated with
specified $\mathrm{p}K$ values over extended or restricted range of
$\mathrm{pH}$ values allows recovery of these same $\mathrm{p}K$ values.
Accepting that $\mathrm{pH} = -1$ would hardly be realized in biological
fluid and, anyway would require dramatic ionic strengths invalidating
conventional understanding of $\mathrm{pH}$ modeling, the consequence
for modeling of assuming the first $\mathrm{p}K_1$ fixed at different
low values will be examined. In so far as the developed model for
$\mathrm{pH}$ dependent charge on HEPES implies a strict ordering of
$\mathrm{p}K$ values, as will be seen in the Methods section, the
consequence for recovery of known $\mathrm{p}K$ values of assuming
strict ordering will be assessed. The final model, thus developed, was
used with the original observed data to obtain best values for the
underlying $\mathrm{p}K$ values and the fit assessed. Finally, the
sensitivity of the final model to imposed noise in $\mathrm{SID}$,
$\mathrm{[HEPES]}$, or $\mathrm{pH}$ will be presented.

Methods
=======

Assuming that $\mathrm{p}K_{a1}$ of the sulfonic group is much lower
than those on the piperazine nitrogens as also supported by conversation
on StackExchange [@ChemSE:HEPES-pK], we will have the following sequence
moving from low to high $\mathrm{pH}$:

```{r echo=FALSE}
# A dash! A dash! My kingdom for a dash!
# (The things that rmarkdown users have to do without
# \usepackage{mhchem})
```
$\mathrm{I: H^{+}N_1\text-HEP\text-N_4H^{+}SO_3H}$ $\sim \mathrm{p}K_{a1}$ on
sulfonic group

$\mathrm{II: H^{+}N_1\text-HEP\text-N_4H^{+}SO_3^{-}}$ $\sim \mathrm{p}K_{a2}$ on
$\mathrm{N_1}$

$\mathrm{III: N_1\text-HEP\text-N_4H^{+}SO_3^{-}}$ $\sim \mathrm{p}K_{a3}$ on
$\mathrm{N_4}$

$\mathrm{IV: N_1\text-HEP\text-N_4SO_3^{-}}$

Now $\mathrm{I + II + III + IV = HEPTOT}$

From the individual dissociations we have

\[
\mathrm{IV} = \frac{
 \mathrm{III} \cdot 10^{-\mathrm{p}K_{a3}}
}{
 [\mathrm{H}^{+}]
} = \frac{
 \mathrm{II} \cdot 10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}}
}{
 [\mathrm{H}^{+}]^2
} = \frac{
 \mathrm{I}
  \cdot 10^{-\mathrm{p}K_{a3}}
  \cdot 10^{-\mathrm{p}K_{a2}}
  \cdot 10^{-\mathrm{p}K_{a1}}
}{
 [\mathrm{H}^{+}]^3
}
\]

\[
\mathrm{I} = \frac{\mathrm{IV} [\mathrm{H}^{+}]^3}{
 10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}} \cdot 10^{-\mathrm{p}K_{a1}}
}
\]

\[
\mathrm{II} = \frac{\mathrm{IV} [\mathrm{H}^{+}]^2}{
 10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}}
}
\]

\[
\mathrm{III} = \frac{\mathrm{IV} [\mathrm{H}^{+}]}{
 10^{-\mathrm{p}K_{a3}}
}
\]

\[
\mathrm{I} + \mathrm{II} + \mathrm{III} = \mathrm{IV} \cdot \left[
 \frac{[\mathrm{H}^{+}]^3}{
  10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}} \cdot 10^{-\mathrm{p}K_{a1}}
 }
 +
 \frac{[\mathrm{H}^{+}]^2}{
  10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}}
 }
 +
 \frac{[\mathrm{H}^{+}]}{
  10^{-\mathrm{p}K_{a3}}
 }
\right] = \mathrm{HEPTOT} - \mathrm{IV}
\]

\[
\mathrm{IV} = \frac{\mathrm{HEPTOT}}{
 1 + \left[
  \frac{[\mathrm{H}^{+}]^3}{
   10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}} \cdot 10^{-\mathrm{p}K_{a1}}
  }
  +
  \frac{[\mathrm{H}^{+}]^2}{
   10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}}
  }
  +
  \frac{[\mathrm{H}^{+}]}{
   10^{-\mathrm{p}K_{a3}}
  }
 \right]
}
\]

So, we can see that $\mathrm{IV}$ is identified from total HEPES
concentration at a given $\mathrm{pH}$ if the three $\mathrm{p}K$ values
are known, and the derivation clearly requires that $\mathrm{p}K_{a1} <
\mathrm{p}K_{a2} < \mathrm{p}K_{a3}$. Consequently, we have quantified
the concentration of all HEPES moieties. Furthermore, the net charge on
each of the four HEPES components are known: $\mathrm{I}: +2$;
$\mathrm{II}: +1$; $\mathrm{III}: 0$; $\mathrm{IV}: -1$. Thereby we have
obtained a specification for the $\mathrm{pH}$ dependent net charge on
HEPES given its total concentration and given the three named
$\mathrm{p}K$ values.  Therefore, with SID denoting strong ion
difference[@Ring2022] and $K_w$ the auto-dissociation of water, the new
charge-balance model for HEPES is:

\begin{equation}
[\mathrm{H}^{+}] + SID - \frac{K_w}{[\mathrm{H}^{+}]} + \mathrm{I} \cdot 2
+ \mathrm{II} - \mathrm{IV} = 0
\end{equation}

With the three $\mathrm{p}K_a$ values for HEPES taken to be $-1$, $3$
and $7.55$ this will be used to estimate $\mathrm{pH}$ in experimental
fluids to compare with observed $\mathrm{pH}$. It should be noted that
Kildeberg[@Kildeberg1983] derives for any multiprotic substance the
average charge given an ordered sequence of $\mathrm{p}K_1 <
\mathrm{p}K_2 \dots < \mathrm{p}K_n$ as
$N = \frac{
 10^{\mathrm{pH} - \mathrm{p}K_1} +
 2 \cdot 10^{2 \mathrm{pH} - \mathrm{p}K_1 - \mathrm{p}K_2} +
 \dots +
 k \cdot  10^{k \mathrm{pH} - \mathrm{p}K_1 \dots \mathrm{p}K_k}
}{
 1+
 10^{\mathrm{pH} - \mathrm{p}K_1} +
 \cdot 10^{2 \mathrm{pH} - \mathrm{p}K_1 - \mathrm{p}K_2} +
 \dots +
 \cdot  10^{k \mathrm{pH} - \mathrm{p}K_1 \dots \mathrm{p}K_k}
} - n$
with $n$ denoting the charge on the fully protonated species. For given
$\mathrm{p}K$ values and $n$ this yields exactly the same charge as
derived above, but since we are also interested in inverse inference
about the $\mathrm{p}K$ values based on observed chemistry and
$\mathrm{pH}$, the first formulation derived here is more malleable when
specifying required order of $\mathrm{p}K$ entries.

Next, using the observed $\mathrm{pH}$ it was attempted to recover the
$\mathrm{p}K_a$ values for HEPES based on known chemical composition.
This included generating _in silico_ data with known $\mathrm{p}K_a$
values for HEPES and recovering again the $\mathrm{p}K_a$ values from
the synthetic data and examining the sensitivity of parameter
estimates/recovery to changing $\mathrm{pH}$ by modification of SID or
by changing HEPES concentration. The $\mathrm{p}K_a$ values for HEPES
were found from non-linear regression using **R** package `nlsr`
[@CRAN:nlsr] and the ordering of the $\mathrm{p}K_a$ values were
enforced by a difference method documented in supplementary materials.

References
==========