### 2023-11-18 manuscript.Rmd
Least-squares solution of charge-balance equations
Instead of solving the charge-balance equation at every known SID, we
can use the least squares method to solve them all together, minimising
the sum of squared change imbalances.

This gives us worse results due to a different error structure of the
model, but may be worth noting.

---
title: Modeling HEPES buffer with a charge-balance approach
author: Troels Ring, Søren Risom Kristensen, Ivan Krylov, John C Nash.
bibliography: manuscript.bib
---

```{r, echo=FALSE}
# NOTE: bookdown features (e.g. cross-references) used below
# This requires the use of a bookdown output format, e.g.
# bookdown::pdf_document2 or bookdown::html_document2
```

Abstract
========

Introduction
============

It has been shown [@Ring2022] that $\mathrm{pH}$ in arbitrary fluids can be
understood based on the three fundamental principles of
electroneutrality, conservation of mass, and rules of dissociation as
devised from physical chemistry. It has also been shown [@Ring2022] that the
charge-balance approach allows definition of the $\mathrm{pH}$-dependent
buffer-capacity, and provides a definition of acid content as titratable
acidity, based on previous work by Kildeberg [@Kildeberg1983]. Recently
it has been reiterated [@Ring2023-charge] that it is absolutely necessary for this
modeling approach to take into account the strong ions, which are those
assumed to be completely dissociated at any $\mathrm{pH}$. In contrast to the
physico-chemical model of Stewart [@Stewart1978], the charge-balance
model (CBM, henceforth) has no separation between independent and
dependent variables. Under the CBM all that can be known about the
system is gathered in the expression of charge in an explicitly stated
chemical system. Consequently, if, as an example, $\mathrm{pH}$ is known and charge
is a known deterministic function of the chemical specification, inverse
inferences on unknown variables can, at least sometimes, be obtained by
root finding, since we know that the charge must be zero. In simple
systems, such as a single monovalent weak acid and strong ions, a single
unknown variable is explicitly estimable [@Ring2022]. The question
focused here is if CBM is rich enough to allow inverse inference in more
complex systems such as with multiprotic buffers. This question arose
under experiments with HEPES buffered sulfate containing fluids
[@Ring2023-acid] in which HEPES was modeled as a monovalent buffer with
pKa 7.5 [@Ring2010] since it was found that estimated $\mathrm{pH}$ was much too
low when observed $\mathrm{pH}$ was less than about 6. HEPES is a zwitterion with
formula 2-[4-(2-Hydroxyethyl)piperazin-1-yl]ethane-1-sulfonic acid and
structure as shown in \@ref(fig:hepes).

```{r hepes, fig.cap = '2-[4-(2-Hydroxyethyl)piperazin-1-yl]ethane-1-sulfonic acid', echo = FALSE}
knitr::include_graphics('HEPES.svg', auto_pdf = TRUE)
```

Wikipedia informs [@WP:HEPES] that HEPES has two pK values, one at 3 and
one at 7.5, but the reference given [@Johnson2014] in Wikipedia states
that "3" comes from the sulfonate group and 7.5 from one of the
piperazine nitrogens.  At $20^\circ\mathrm{C}$ Good et al [@Good1966]
has pKa 3 and 7.55 without detailing which groups are at work, while
another study [@AlGhobashy2014] mentions that 3 and 7.5 both belongs to
piperazine amines and informs that the sulfonate group is dissociated
over "almost the whole $\mathrm{pH}$ range".  Furthermore, is has been stated that
the nitrogen at the $\mathrm{OH}$ end of HEPES is protonated in methanol
and the nitrogen in the sulfonate end is protonated in water
[@Sledz2010].

Hence it was not clear how best to model the $\mathrm{pH}$-dependent
charge on HEPES to estimate $\mathrm{pH}$ based on known chemistry.
Furthermore, it was desired to make clear if the $\mathrm{p}K$ values
for HEPES could be elucidated based on measured $\mathrm{pH}$. Herein,
these questions will be approached by first developing a model for the
mean charge on HEPES as a function of $\mathrm{pH}$ and applying this to
proposed values of set of underlying pKs using data obtained previously
[@Ring2023-acid]. Next, it will be examined if data generated with
specified $\mathrm{p}K$ values over extended or restricted range of
$\mathrm{pH}$ values allows recovery of these same $\mathrm{p}K$ values.
Accepting that $\mathrm{pH} = -1$ would hardly be realized in biological
fluid and, anyway would require dramatic ionic strengths invalidating
conventional understanding of $\mathrm{pH}$ modeling, the consequence
for modeling of assuming the first $\mathrm{p}K_1$ fixed at different
low values will be examined. In so far as the developed model for
$\mathrm{pH}$ dependent charge on HEPES implies a strict ordering of
$\mathrm{p}K$ values, as will be seen in the Methods section, the
consequence for recovery of known $\mathrm{p}K$ values of assuming
strict ordering will be assessed. The final model, thus developed, was
used with the original observed data to obtain best values for the
underlying $\mathrm{p}K$ values and the fit assessed. Finally, the
sensitivity of the final model to imposed noise in $\mathrm{SID}$,
$\mathrm{[HEPES]}$, or $\mathrm{pH}$ will be presented.

Methods
=======

Assuming that $\mathrm{p}K_{a1}$ of the sulfonic group is much lower
than those on the piperazine nitrogens as also supported by conversation
on StackExchange [@ChemSE:HEPES-pK], we will have the following sequence
moving from low to high $\mathrm{pH}$:

```{r echo=FALSE}
# A dash! A dash! My kingdom for a dash!
# (The things that rmarkdown users have to do without
# \usepackage{mhchem})
```
$\mathrm{I: H^{+}N_1\text-HEP\text-N_4H^{+}SO_3H}$ $\sim \mathrm{p}K_{a1}$ on
sulfonic group

$\mathrm{II: H^{+}N_1\text-HEP\text-N_4H^{+}SO_3^{-}}$ $\sim \mathrm{p}K_{a2}$ on
$\mathrm{N_1}$

$\mathrm{III: N_1\text-HEP\text-N_4H^{+}SO_3^{-}}$ $\sim \mathrm{p}K_{a3}$ on
$\mathrm{N_4}$

$\mathrm{IV: N_1\text-HEP\text-N_4SO_3^{-}}$

Now $\mathrm{I + II + III + IV = HEPTOT}$

From the individual dissociations we have

\[
\mathrm{IV} = \frac{
 \mathrm{III} \cdot 10^{-\mathrm{p}K_{a3}}
}{
 [\mathrm{H}^{+}]
} = \frac{
 \mathrm{II} \cdot 10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}}
}{
 [\mathrm{H}^{+}]^2
} = \frac{
 \mathrm{I}
  \cdot 10^{-\mathrm{p}K_{a3}}
  \cdot 10^{-\mathrm{p}K_{a2}}
  \cdot 10^{-\mathrm{p}K_{a1}}
}{
 [\mathrm{H}^{+}]^3
}
\]

\[
\mathrm{I} = \frac{\mathrm{IV} [\mathrm{H}^{+}]^3}{
 10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}} \cdot 10^{-\mathrm{p}K_{a1}}
}
\]

\[
\mathrm{II} = \frac{\mathrm{IV} [\mathrm{H}^{+}]^2}{
 10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}}
}
\]

\[
\mathrm{III} = \frac{\mathrm{IV} [\mathrm{H}^{+}]}{
 10^{-\mathrm{p}K_{a3}}
}
\]

\[
\mathrm{I} + \mathrm{II} + \mathrm{III} = \mathrm{IV} \cdot \left[
 \frac{[\mathrm{H}^{+}]^3}{
  10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}} \cdot 10^{-\mathrm{p}K_{a1}}
 }
 +
 \frac{[\mathrm{H}^{+}]^2}{
  10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}}
 }
 +
 \frac{[\mathrm{H}^{+}]}{
  10^{-\mathrm{p}K_{a3}}
 }
\right] = \mathrm{HEPTOT} - \mathrm{IV}
\]

\[
\mathrm{IV} = \frac{\mathrm{HEPTOT}}{
 1 + \left[
  \frac{[\mathrm{H}^{+}]^3}{
   10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}} \cdot 10^{-\mathrm{p}K_{a1}}
  }
  +
  \frac{[\mathrm{H}^{+}]^2}{
   10^{-\mathrm{p}K_{a3}} \cdot 10^{-\mathrm{p}K_{a2}}
  }
  +
  \frac{[\mathrm{H}^{+}]}{
   10^{-\mathrm{p}K_{a3}}
  }
 \right]
}
\]

So, we can see that $\mathrm{IV}$ is identified from total HEPES
concentration at a given $\mathrm{pH}$ if the three $\mathrm{p}K$ values
are known, and the derivation clearly requires that $\mathrm{p}K_{a1} <
\mathrm{p}K_{a2} < \mathrm{p}K_{a3}$. Consequently, we have quantified
the concentration of all HEPES moieties. Furthermore, the net charge on
each of the four HEPES components are known: $\mathrm{I}: +2$;
$\mathrm{II}: +1$; $\mathrm{III}: 0$; $\mathrm{IV}: -1$. Thereby we have
obtained a specification for the $\mathrm{pH}$ dependent net charge on
HEPES given its total concentration and given the three named
$\mathrm{p}K$ values.  Therefore, with SID denoting strong ion
difference[@Ring2022] and $K_w$ the auto-dissociation of water, the new
charge-balance model for HEPES is:

\begin{equation}
[\mathrm{H}^{+}] + \mathrm{SID} - \frac{K_w}{[\mathrm{H}^{+}]}
+ \mathrm{I} \cdot 2 + \mathrm{II} - \mathrm{IV} = 0
(\#eq:charge-balance)
\end{equation}

With the three $\mathrm{p}K_a$ values for HEPES taken to be $-1$, $3$
and $7.55$ this will be used to estimate $\mathrm{pH}$ in experimental
fluids to compare with observed $\mathrm{pH}$. It should be noted that
Kildeberg[@Kildeberg1983] derives for any multiprotic substance the
average charge given an ordered sequence of $\mathrm{p}K_1 <
\mathrm{p}K_2 \dots < \mathrm{p}K_n$ as
$N = \frac{
 10^{\mathrm{pH} - \mathrm{p}K_1} +
 2 \cdot 10^{2 \mathrm{pH} - \mathrm{p}K_1 - \mathrm{p}K_2} +
 \dots +
 k \cdot  10^{k \mathrm{pH} - \mathrm{p}K_1 \dots \mathrm{p}K_k}
}{
 1+
 10^{\mathrm{pH} - \mathrm{p}K_1} +
 \cdot 10^{2 \mathrm{pH} - \mathrm{p}K_1 - \mathrm{p}K_2} +
 \dots +
 \cdot  10^{k \mathrm{pH} - \mathrm{p}K_1 \dots \mathrm{p}K_k}
} - n$
with $n$ denoting the charge on the fully protonated species. For given
$\mathrm{p}K$ values and $n$ this yields exactly the same charge as
derived above, but since we are also interested in inverse inference
about the $\mathrm{p}K$ values based on observed chemistry and
$\mathrm{pH}$, the first formulation derived here is more malleable when
specifying required order of $\mathrm{p}K$ entries.

Next, using the observed $\mathrm{pH}$ it was attempted to recover the
$\mathrm{p}K_a$ values for HEPES based on known chemical composition.
This included generating _in silico_ data with known $\mathrm{p}K_a$
values for HEPES and recovering again the $\mathrm{p}K_a$ values from
the synthetic data and examining the sensitivity of parameter
estimates/recovery to changing $\mathrm{pH}$ by modification of SID or
by changing HEPES concentration. The $\mathrm{p}K_a$ values for HEPES
were found from non-linear regression using **R** package `nlsr`
[@CRAN:nlsr] and the ordering of the $\mathrm{p}K_a$ values were
enforced by a difference method documented in supplementary materials.

Appendix: Software considerations
=================================

The `nlsr` package @CRAN:nlsr which has been used for the estimation of
parameters in the nonlinear models for this article has some features that 
more easily permitted the computations. Moreover, discussions concerning how
to use the software led to unexpected possibilities in that use.

### Ensuring parameters follow a specific ordering

We require that the three $\mathrm{p}K_a$ values be in increasing order of
numerical value. To simplify the discussion here, we will write these as 
`p1`, `p2` and `p3` within our R program scripts. Thus we want

```
       p1  <  p2  < p3
```

One approach, which we have used here, is to reparameterize the formulas for
the model so

``` 
      p2  =  p1 + d12
      p3  =  p1 + d12 + d23
```

If we constrain `d12` and `d23` to be positive, or at least non-negative, then
the new parameterization will ensure the ordering. However, this does require 
that the software can offer bounds constraints on the parameters. `nslr` is 
capable in this way, and even allows upper and lower bounds to be equal, 
resulting in fixed parameters, or **masks**. Other software in R has similar
capabilities, but the built-in `nls()` function only handles bounds if the
`port` algorithm is used, and there are warnings that this code may be 
incomplete. Package `minpack.lm` also handles bounds, though a failure was
encountered in one example (@NashBhattacharjee).

### Vectorizing a function

**IK**: This feels like I'm writing a tutorial. Should it be less
detailed?

**R** scripts naturally lend themselves to vectorization, thanks to the
language performing many operations on whole vectors. For example, **R**
has no problem evaluating the charge-balance equation function for a
vector of different values of $[\mathrm H]$ with other values remaining
scalars:

```{r HEPESFUNC}
kw <- 1e-14
HEPESFUNC <- function(H,SID,HEPTOT,pK1,pK2,pK3) {
 XX <- (H^3/(10^-pK3*10^-pK2*10^-pK1)+H^2/(10^-pK3*10^-pK2)+H/(10^-pK3))
 IV <- HEPTOT/(1+XX)
 I <- IV*H^3/(10^-pK3*10^-pK2*10^-pK1)
 II <- IV*H^2/(10^-pK3*10^-pK2)
 III <- IV*H/10^-pK3
 H - kw/H + SID + I*2 + II - IV
}
```

This principle is used by the non-linear regression functions which, when
given a formula `y ~ f(x, theta)` and vectors `x` of independent
variables and `y` of observed values of the dependent variable, call the
model function `f` with the whole vector `x` as argument, expecting to
receive an estimate of the whole `y` at the current value of the
parameter `theta` in response.

Some functions are not naturally vectorized and cannot be easily changed
to obtain this property. For example, the model function for a
non-linear regression problem may involve root-finding using `uniroot`:

```{r}
# Which pH value satisfies the charge-balance equation?
pH <- function(SID, HEPTOT, pK1, pK2, pK3)
  -log10(uniroot(HEPESFUNC,c(1e-20,1),tol=1e-20,maxiter=1e4,
                 HEPTOT=HEPTOT,SID = SID, pK1=pK1,pK2=pK2,pK3=pK3)$root)
```

The underlying mechanism of `uniroot()` only works with scalar functions
of one argument, so passing vectors of length more than `1` to this
function will cause an error: `uniroot()` is not prepared to find
multiple different roots for every given set of the parameter values. In
order to use this function with non-linear regression methods, a wrapper
is needed. This wrapper must call the original function in a loop for
every element of the vector arguments passed to it:

```{r}
pH.wrapper <- function(SID, HEPTOT, pK1, pK2, pK3)
 mapply(function(SID, HEPTOT)
  -log10(uniroot(HEPESFUNC, c(1e-20,1), tol = 1e-20, maxiter = 1e4,
                 HEPTOT = HEPTOT, SID = SID,
                 pK1 = pK1, pK2 = pK2, pK3 = pK3)$root),
  SID = SID, HEPTOT = HEPTOT
 )
```

The base function `Vectorize` automates this transformation. It returns
a wrapper function that obtains the values of every argument given to
the wrapper and passes them to a live-constructed call to `mapply()`,
letting the latter call the original function. A call to
`mapply(f, a, b, c)`, returns a vector of values of
`f(a[i], b[i], c[i])` for every applicable `i`, subject to vector
recycling rules:

```{r}
ans <- c(
 pH(.04, .05, -1, 3, 7.55),
 pH(.035, .05, -1, 3, 7.55)
)
all.equal(ans, pH.wrapper(c(.04, .035), .05, -1, 3, 7.55))
pHm <- Vectorize(pH)
all.equal(ans, pHm(c(.04, .035), .05, -1, 3, 7.55))
```

Unfortunately, the use of automatic vectorization precludes the use of
fully automatic symbolic differentiation: static code analysis will not
go past `vapply`, `mapply`, or `uniroot`. Still, it is possible to
construct most of the expression piece by piece and have it
differentiated this way:

```{r}
charge_balance <- local({
	XX <- quote(H^3/(10^-pK3*10^-pK2*10^-pK1)+H^2/(10^-pK3*10^-pK2)+H/(10^-pK3))
	IV <- substitute(HEPTOT/(1+XX))
	I <- substitute(IV*H^3/(10^-pK3*10^-pK2*10^-pK1))
	II <- substitute(IV*H^2/(10^-pK3*10^-pK2))
	III <- substitute(IV*H/10^-pK3)
	nlsr::codeDeriv(
		substitute(H - kw/H + SID + I*2 + II - IV),
		c('H', 'pK1', 'pK2', 'pK3')
	)
})
# a few screenfuls of code; >2500 pages of output with verbose=TRUE
```

In order to obtain the final Jacobian, we will have to somehow
differentiate past the root-finding operation. For that we will use the
implicit function theorem. Let the charge-balance equation be
$f([\mathrm H], \theta) = 0$, with $[\mathrm H]$ being the desired
equilibrium concentration of protons and $\theta$ symbolising all of the
parameters of the equation. It implicitly defines the function
$[\mathrm H]_\mathrm{eq}(\theta)$ that satisfies
$F(\theta) = f([\mathrm H]_\mathrm{eq}(\theta), \theta) = 0$. Differentiating the
latter equation by $\theta$, we get $\frac{dF}{d\theta} =
\frac{\partial{}f}{\partial[\mathrm H]}
\frac{d[\mathrm H]_\mathrm{eq}}{d\theta} +
\frac{\partial{}f}{\partial\theta} \frac{d\theta}{d\theta} = 0$, from
which we get $\frac{d[\mathrm H]_\mathrm{eq}}{d\theta} =
-\frac{\partial{}f/\partial\theta}{\partial{}f/\partial[\mathrm H]}$:

```{r}
pH.sym <- function(SID, HEPTOT, pK1, pK2, pK3) {
 H <- vapply(SID, function(SID) { # for every point:
  # find the root of the function
  H <- uniroot(HEPESFUNC, c(1e-20,1), tol = 1e-20, maxiter = 1e4,
                 HEPTOT = HEPTOT, SID = SID,
                 pK1 = pK1, pK2 = pK2, pK3 = pK3)$root
  # evaluate the derivatives by H and parameters
  g <- attr(eval(charge_balance), 'gradient')
  # and calculate the derivative of the H_eq by parameters;
  # combine together for easier access by vapply
  c(H, -g[,2:4]/g[,1])
 }, numeric(4))
 # now take the negative logarithm to obtain pH
 structure(-log10(H[1,]), gradient = -(1/(H[1,] * log(10))) * t(H[-1,]))
}
pH.sym(c(.04, .035), .05, -1, 3, 7.55)
```

This gives similar, but not exactly the same results as numeric
differentiation:

```{r}
nlsr::numericDerivR(
 quote(pHm(c(.04, .035), .05, pK1, pK2, pK3)),
 c('pK1', 'pK2', 'pK3'),
 list2env(list(pK1 = -1, pK2 = 3, pK3 = 7.55)),
 central = TRUE
)
```

### Solving the charge-balance equations in the least squares sense

Yet another way to model the data is to plug the experimental values of
$\mathrm{SID}$ and measured $\mathrm{pH}$ straight into the
charge-balance equation \@ref(eq:charge-balance), without going through
the intermediate step of finding the negative logaritm of the root of
the function. This changes the problem: instead of
$\min_{\mathrm{p}K_{a,1}, \mathrm{p}K_{a,2}, \mathrm{p}K_{a,3}}
\sum_i(\mathrm p\hat{\mathrm H}(\mathrm{SID}_i, \mathrm{p}K_{a,1},
\mathrm{p}K_{a,2}, \mathrm{p}K_{a,3}) - \mathrm{pH}_i)^2$, we are now
minimising the sum of squared residuals from eq.
\@ref(eq:charge-balance) itself:

```{r}
load('HEPES.RData') # provides the DD data.frame
charge_balance_fn <- function() NULL
formals(charge_balance_fn) <- formals(HEPESFUNC)
body(charge_balance_fn) <- local({
	XX <- quote(H^3/(10^-pK3*10^-pK2*10^-pK1)+H^2/(10^-pK3*10^-pK2)+H/(10^-pK3))
	IV <- substitute(HEPTOT/(1+XX))
	I <- substitute(IV*H^3/(10^-pK3*10^-pK2*10^-pK1))
	II <- substitute(IV*H^2/(10^-pK3*10^-pK2))
	III <- substitute(IV*H/10^-pK3)
	nlsr::codeDeriv(
		substitute(H - kw/H + SID + I*2 + II - IV),
		c('pK1', 'pK2', 'pK3')
	)
})

loss <- function(x) charge_balance_fn(
	H = 10^-DD$OBS, SID = DD$SID, HEPTOT = DD$HEPES,
	pK1 = x['pK1'], pK2 = x['pK2'], pK3 = x['pK3']
)
(m_cb <- nlsr::nlfb(
	start = c(pK1 = -1, pK2 = 3, pK3 = 7.55),
	resfn = loss, jacfn = loss,
	lower = -3, upper = 9
))
```

While this avoids finding the roots of the charge-balance equation, the
implied residual structure (normally-distributed errors in the
$\mathrm{pH}$ measurements in the original model vs.
normally-distributed charge imbalances in concentration units in the
current model) is different, and so are the resulting estimates:

```{r}
(m_rf <- nlsr::nlxb(
	OBS ~ pH.wrapper(SID, HEPES, pK1, pK2, pK3),
	data = DD, start = c(pK1 = -1, pK2 = 3, pK3 = 7.55),
	lower = -3, upper = 9,
	control = list(japprox = 'jacentral')
))
```

### Gauging the reliability of the estimated parameters

All parameter estimation methods must contend with the possibility that 
a variety of parameter values may give very similar magnitudes for the loss
function (e.g, sum of squared residuals) that is optimized to obtain the
estimates. In the case of estimation that can be converted to minimizing a
sum of squares, the Jacobian of the vector of residuals will become close 
to singular if parameters are indeterminate. This can be judged by computing
the **singular values** (??JN: do we want a reference) of the Jacobian and
package `nlsr` does this and displays the values when a summary of a solution 
is requested. A rule of thumb is to regard the Jacobian as essentially 
singular if the ratio of the smallest to largest singular value is "small".
A value of smaller than 1e-4 would be a typical criterion. Note that more than
one singular value may be small, indicating that there is more than one 
dimension of indeterminacy. However, such indicators apply to the ensemble of
parameters, and as such concern the overall model.

References
==========