### 2023-11-30 noise.Rmd
Draw the plots from the noise experiment

First let's load the results produced by `noise.R` overnight.

The results are organised a bit clumsily, with `pK_true` being a list of
numeric vectors. Our goal is to group by every experiment condition (set
of true $\mathrm{p}K_a$ values, number of points, noise-to-signal ratio,
kind of noise) and summarise the $10$ repeats that were conducted with
different samples of noise.

```{r}
n <- readRDS('noise.rds')
# On its own, `by` cannot group on list columns, so produce a surrogate
# key for the true pK column, representing numeric vectors as strings.
# This is probably not efficient at all, but will work for our purposes.
n$idx <- as.factor(sapply(n$pK_true, paste, collapse = ' '))
levels(n$idx) <- seq_along(levels(n$idx))
pn <- do.call(rbind, by(
	n, n[c('noise_ratio','n_points','noise_kind', 'idx')],
	function(d) {
		pK_true <- d$pK_true[[1]]
		# combine all estimated pKs into one matrix
		pK_est <- simplify2array(d$pK_est)
		d$pK_est <- NULL
		cbind(d[1,],
			pK_est = I(list(t(pK_est))),
			# average residual: should approach 0 for unbiased estimator
			res_mean = mean(pK_est - pK_true),
			# deviance of residuals: measures the spread of the
			# estimates
			res_dev = mean((pK_est - pK_true)^2)
		)
	}
))
```

Every row of the newly obtained `data.frame` now contains results for
every repeat of the same set of experimental conditions:

```{r}
head(pn)
pn[[1,'pK_est']]
```

Let's reduce it further, averaging every generated set of true
$\mathrm{p}K_a$ values:

```{r}
pna <- aggregate(
	pn[c('res_mean','res_dev')],
	pn[c('noise_ratio','n_points','noise_kind')],
	mean
)
```

This lets us plot the average estimation error and the standard
deviation of the estimates:
```{r}
lattice::levelplot(
	res_mean ~ noise_ratio + n_points | noise_kind, pna,
	group = noise_kind, scales = list(log = 10),
	main = quote(
		'Average '*pK[a]*' estimation error '*
		'(should be close to 0)'
	),
	xlab = 'Noise-to-signal ratio',
	ylab = 'Number of known points',
	cuts = 255, col.regions = hcl.colors(256, 'Broc')
)
```

I don't see any pattern in the data, and it seems to be fairly centered
around $0$, so the estimate is effectively unbiased, even in presence of
non-normally-distributed noise. But what about the variance?
```{r}
lattice::levelplot(
	sqrt(res_dev) ~ noise_ratio + n_points | noise_kind, pna,
	group = noise_kind, scales = list(log = 10),
	main = quote(
		'Standard deviation of '*pK[a]*' estimate'
	),
	xlab = 'Noise-to-signal ratio',
	ylab = 'Number of known points',
	cuts = 255, col.regions = hcl.colors(256, 'Terrain 2')
)
```

Given $10^{1.5}$ points and realistic (?) noise, the standard deviation
of the $\mathrm{p}K_a$ estimates can be as high as $1.5$. Under
three-sigma rule, this gives us an error of up to $4.5$ $\mathrm{pH}$
units...